\documentclass[sigconf,natbib=true,anonymous=true]{acmart}

\usepackage{enumitem}
\setlist{leftmargin=10pt}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,decorations.pathreplacing}

\AtBeginDocument{%
  \providecommand\BibTeX{{Bib\TeX}}}

% Placeholder for camera-ready
\setcopyright{none}

\newcommand{\topdown}{\textsc{TopDown}}
\newcommand{\bottomup}{\textsc{BottomUp}}
\newcommand{\dualend}{\textsc{DualEnd}}
\newcommand{\bidir}{\textsc{BiDir}}
\newcommand{\cocktail}{\textsc{CocktailSort}}
\newcommand{\deselection}{\textsc{DESelection}}
\begin{document}

\title{Finding the Worst Is Harder Than Finding the Best: Directional Asymmetry in LLM Setwise Ranking}

\author{Anonymous}

\begin{abstract}
Large Language Models (LLMs) have emerged as effective zero-shot document rerankers. Among the four prompting paradigms---pointwise, pairwise, listwise, and setwise---the setwise approach offers a compelling trade-off between effectiveness and efficiency by asking the LLM to select the most relevant document from a candidate set. However, current setwise methods only extract one piece of ranking information per LLM call: the \emph{best} document. In this paper, we investigate bidirectional selection strategies that extract more information from each comparison. We propose three novel approaches: (1) \textbf{Bottom-Up Setwise Ranking}, which reverses the selection to identify the \emph{least} relevant document, building rankings from the bottom up; (2) \textbf{Dual-End Setwise Ranking}, which asks the LLM to simultaneously identify both the most and least relevant documents, extracting twice the ranking information per call; and (3) \textbf{Bidirectional Ensemble}, which fuses top-down and bottom-up rankings. We implement these strategies with two new sorting algorithms---cocktail shaker sort and double-ended selection sort---designed for dual-end comparisons. Experiments on TREC DL 2019/2020 with nine LLMs spanning three architecture families (Flan-T5, Qwen3, and Qwen3.5, ranging from 780M to 27B parameters) reveal three key findings. First, the \dualend{} family is the strongest overall: \dualend{}-Cocktail achieves the best average NDCG@10 on both datasets, while DualEnd methods (Cocktail or Selection) achieve the highest NDCG@10 in 14 of 18 model--dataset configurations and improve over the best top-down baseline by up to 4.5 points. Second, \bottomup{} is consistently weaker than top-down, refuting the hypothesis that LLMs are better at identifying irrelevant documents. Third, bidirectional ensemble does \emph{not} outperform the top-down baseline, indicating that directional complementarity alone is insufficient. Per-query significance tests are more conservative: only one DualEnd gain remains Bonferroni-significant, while the negative BottomUp and BiDir results are more robust. Our information-theoretic analysis shows that dual-end extracts $\sim$79\% more ranking information per LLM call at negligible prompt overhead, helping explain why joint best-worst elicitation is a useful quality-first refinement despite its cost.
\end{abstract}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002951.10003317.10003347.10003356</concept_id>
       <concept_desc>Information systems~Rank aggregation</concept_desc>
       <concept_significance>500</concept_significance>
   </concept>
   <concept>
       <concept_id>10002951.10003317.10003338</concept_id>
       <concept_desc>Information systems~Retrieval models and ranking</concept_desc>
       <concept_significance>500</concept_significance>
   </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Rank aggregation}
\ccsdesc[500]{Information systems~Retrieval models and ranking}

\keywords{LLM-based ranking, setwise prompting, zero-shot reranking, document ranking}

\maketitle

% =============================================================================
\section{Introduction}
\label{sec:intro}
% =============================================================================

Large Language Models (LLMs) have demonstrated remarkable capabilities as zero-shot passage rerankers in information retrieval (IR) systems~\citep{sun2023chatgpt, qin2024large, zhuang2024setwise}. Given a query and a set of candidate documents retrieved by a first-stage retriever, LLMs can reorder candidates by relevance without any task-specific fine-tuning. Four prompting paradigms have been established: \emph{pointwise} (score each document independently), \emph{pairwise} (compare documents in pairs), \emph{listwise} (rerank a full list), and \emph{setwise} (select the best from a small candidate set)~\citep{zhuang2024setwise}.

The \textbf{setwise approach}~\citep{zhuang2024setwise} occupies a sweet spot in the effectiveness--efficiency trade-off. By presenting a small set of candidates (e.g., 4 documents) and asking the LLM ``\emph{which passage is the most relevant?}'', it extracts more comparative information than pairwise (which compares only 2) while requiring shorter prompts than listwise (which presents 10--20). Combined with classical sorting algorithms like heapsort and bubblesort, setwise ranking achieves high effectiveness with significantly fewer LLM calls than pairwise approaches.

However, current setwise methods have a fundamental limitation: \textbf{each LLM call extracts only one ranking decision}---the identity of the \emph{most relevant} document. The comparative assessment the LLM performs internally necessarily evaluates all candidates, yet we discard all information except the winner. This is analogous to running a tournament but only recording gold medals while throwing away all other placement information.

Consider a concrete example: when an LLM is presented with four passages and asked ``which is the most relevant?'', it must internally assess all four before producing an answer. The model effectively \emph{knows} which passage is the worst too---it simply is not asked to report it. From an information-theoretic perspective, selecting the best from $m$ candidates yields $\log_2(m)$ bits of information. But by asking for \emph{both} the best and worst, we obtain $\log_2(m) + \log_2(m-1)$ bits---nearly doubling the information extracted per call with negligible prompt overhead.

In this paper, we ask: \emph{can we extract more ranking information from each LLM comparison?} We investigate three bidirectional strategies:

\begin{enumerate}[leftmargin=*]
\item \textbf{Bottom-Up Setwise Ranking (\bottomup{})}: We reverse the prompt to ask ``\emph{which passage is the least relevant?}'' and build the ranking from the bottom up by iteratively eliminating the worst documents. This tests whether LLMs exhibit different accuracy and bias patterns when identifying irrelevant versus relevant documents.

\item \textbf{Dual-End Setwise Ranking (\dualend{})}: We modify the prompt to ask for \emph{both} the most and least relevant documents simultaneously: ``\emph{which is the most relevant and which is the least relevant?}'' This extracts two ranking decisions per LLM call, potentially halving the number of required comparisons.

\item \textbf{Bidirectional Ensemble (\bidir{})}: We run both top-down (standard) and bottom-up ranking independently and fuse the results using rank aggregation methods. This tests whether the two directions provide complementary relevance signals.
\end{enumerate}

To support dual-end comparisons algorithmically, we introduce two new sorting algorithms adapted for LLM-based ranking:
\begin{itemize}[leftmargin=*]
\item \textbf{Cocktail Shaker Sort}: A bidirectional bubblesort that alternates between forward passes (sinking worst documents down) and backward passes (bubbling best documents up), using dual-end comparisons throughout.
\item \textbf{Double-Ended Selection Sort}: A tournament-based approach that extracts both the global best and global worst from the unsorted pool in each round, filling the ranking from both ends simultaneously.
\end{itemize}

Figure~\ref{fig:overview} illustrates the key differences between standard top-down, bottom-up, and dual-end setwise ranking.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    doc/.style={rectangle, draw, minimum width=0.6cm, minimum height=0.5cm, font=\scriptsize},
    arrow/.style={-{Stealth[length=2mm]}, thick},
    label/.style={font=\small\bfseries},
    >=Stealth
]
% TopDown
\node[label] at (0, 2.6) {\topdown{}};
\node[doc, fill=green!30] (ta) at (-0.9, 2.0) {A};
\node[doc] (tb) at (-0.3, 2.0) {B};
\node[doc] (tc) at (0.3, 2.0) {C};
\node[doc] (td) at (0.9, 2.0) {D};
\node[font=\scriptsize, text=green!50!black] at (0, 1.3) {``Which is \textbf{best}?'' $\to$ A};
\draw[arrow, green!50!black] (0, 1.1) -- (0, 0.7);
\node[doc, fill=green!30] at (-0.45, 0.4) {\scriptsize 1st};
\node[doc, fill=gray!15] at (0.15, 0.4) {\scriptsize ?};
\node[doc, fill=gray!15] at (0.75, 0.4) {\scriptsize ?};

% BottomUp
\node[label] at (3.4, 2.6) {\bottomup{}};
\node[doc] (ba) at (2.5, 2.0) {A};
\node[doc] (bb) at (3.1, 2.0) {B};
\node[doc] (bc) at (3.7, 2.0) {C};
\node[doc, fill=red!30] (bd) at (4.3, 2.0) {D};
\node[font=\scriptsize, text=red!50!black] at (3.4, 1.3) {``Which is \textbf{worst}?'' $\to$ D};
\draw[arrow, red!50!black] (3.4, 1.1) -- (3.4, 0.7);
\node[doc, fill=gray!15] at (2.65, 0.4) {\scriptsize ?};
\node[doc, fill=gray!15] at (3.25, 0.4) {\scriptsize ?};
\node[doc, fill=red!30] at (3.85, 0.4) {\scriptsize last};

% DualEnd
\node[label] at (6.8, 2.6) {\dualend{}};
\node[doc, fill=green!30] (da) at (5.9, 2.0) {A};
\node[doc] (db) at (6.5, 2.0) {B};
\node[doc] (dc) at (7.1, 2.0) {C};
\node[doc, fill=red!30] (dd) at (7.7, 2.0) {D};
\node[font=\scriptsize, text=blue!50!black, align=center] at (6.8, 1.25) {``\textbf{Best} \& \textbf{worst}?''\\$\to$ A, D};
\draw[arrow, blue!50!black] (6.8, 0.75) -- (6.8, 0.55);
\node[doc, fill=green!30] at (6.05, 0.25) {\scriptsize 1st};
\node[doc, fill=gray!15] at (6.65, 0.25) {\scriptsize ?};
\node[doc, fill=red!30] at (7.25, 0.25) {\scriptsize last};
\end{tikzpicture}
\caption{Comparison of setwise ranking strategies. \topdown{} identifies only the best document per call. \bottomup{} identifies only the worst. \dualend{} identifies both, extracting 2$\times$ ranking information per LLM call.}
\label{fig:overview}
\end{figure}

Our contributions are:
\begin{enumerate}[leftmargin=*]
\item We propose three novel bidirectional strategies for LLM-based setwise ranking that extract more information per LLM call (\S\ref{sec:method}).
\item We introduce two sorting algorithms---cocktail shaker sort and double-ended selection sort---designed for dual-end LLM comparisons (\S\ref{sec:algorithms}).
\item We provide formal complexity analysis comparing all methods, including an information-theoretic framework for quantifying ranking information per LLM call (\S\ref{sec:algorithms}).
\item We conduct comprehensive experiments on TREC DL 2019/2020 with nine LLMs across three architecture families (Flan-T5, Qwen3, and Qwen3.5, 780M--27B), plus a representative six-dataset BEIR generalization study on Flan-T5-XL, Qwen3-8B, and Qwen3.5-9B, analyzing effectiveness, efficiency, passage length sensitivity, position bias, and ranking complementarity (\S\ref{sec:results}).
\item We provide an open-source implementation integrated into the \texttt{llm-rankers} framework\footnote{Code available at: \url{https://anonymous}} (\S\ref{sec:experiments}).
\end{enumerate}


% =============================================================================
\section{Related Work}
\label{sec:related}
% =============================================================================

\subsection{LLM-Based Document Ranking}

LLM-based document ranking has rapidly evolved since the introduction of RankGPT~\citep{sun2023chatgpt}, which demonstrated that properly instructed LLMs can deliver competitive results to supervised rerankers. The field has converged on four prompting paradigms:

\textbf{Pointwise methods}~\citep{nogueira2020document, sachan2022improving} score each document independently. MonoT5~\citep{nogueira2020document} uses a sequence-to-sequence model to predict a ``true'' or ``false'' relevance label for each query-document pair, using the probability of the ``true'' token as a relevance score. While highly efficient---each document requires only one inference call and calls are easily parallelizable---pointwise methods suffer from poor ranking effectiveness because they cannot capture relative relevance between documents. A document that is moderately relevant to one query may receive the same score as one that is highly relevant to another, making cross-document comparison unreliable.

\textbf{Pairwise methods}~\citep{qin2024large, pradeep2021expando} compare two documents at a time, asking the LLM which is more relevant. PRP~\citep{qin2024large} achieves competitive effectiveness with GPT-4-class performance using smaller models like Flan-UL2 (20B), outperforming ChatGPT-based pointwise solutions by 4.2\% on NDCG@10. The DuoT5~\citep{pradeep2021expando} model extends the Expando-Mono-Duo pipeline with a T5-based pairwise reranker. While pairwise methods achieve the highest effectiveness due to direct inter-document comparison, they require $O(n^2)$ comparisons for all-pairs or $O(n \log n)$ with sorting algorithms---a significant computational overhead.

\textbf{Listwise methods}~\citep{sun2023chatgpt, ma2023zero} present multiple documents and request a complete ranking permutation. RankGPT~\citep{sun2023chatgpt} uses a sliding window strategy: a window of $w$ documents slides from the bottom of the initial ranking to the top, with the LLM reranking each window. This approach captures rich inter-document relationships and produces coherent rankings. However, it is constrained by context length, sensitive to input order (position bias), and expensive due to long prompts. Rank-without-GPT~\citep{zhang2025rankwithoutgpt} demonstrates that effective listwise rerankers can be built on open-source LLMs, achieving 97\% of GPT-4 effectiveness.

\textbf{Setwise methods}~\citep{zhuang2024setwise} ask the LLM to select the most relevant document from a small candidate set (typically 3--4 documents), using the selection as a comparator in sorting algorithms. This paradigm balances effectiveness and efficiency: it extracts more comparative information per call than pairwise (multiple documents compared simultaneously) while using shorter prompts than listwise (only a small subset presented). The approach has been extended by setwise insertion~\citep{podolak2025beyond}, which leverages the initial BM25 ranking as prior knowledge for 31\% time reduction; tournament-based ranking (TourRank)~\citep{chen2025tourrank}, which uses sports tournament brackets for more robust rankings; and reasoning-enhanced variants (Rank-R1)~\citep{zhuang2025rankr1}, which integrate chain-of-thought reasoning via reinforcement learning.

All existing setwise methods share a common design choice: they only ask the LLM to identify the \emph{most relevant} document. Our work is the first to investigate alternative selection targets (least relevant) and simultaneous multi-target selection (both most and least).

\subsection{Sorting Algorithms for LLM Ranking}

Classical sorting algorithms have been adapted as the backbone for pairwise and setwise LLM ranking. The choice of algorithm significantly impacts both the number of LLM calls and the quality of the final ranking:

\textbf{Heapsort} builds an $n$-ary max-heap where each node is compared against its $c$ children (for setwise with \texttt{num\_child}$=c$). Building the heap requires $O(n)$ comparisons, and extracting $k$ maximums requires $O(k \log_c n)$ comparisons, yielding $O(n + k \log_c n)$ total. For the typical setting ($k=10$, $n=100$, $c=3$), this is approximately 130 comparisons.

\textbf{Bubblesort} starts from the bottom of the list and ``bubbles'' the most relevant document to the top through a series of windowed comparisons. Each of $k$ passes requires $O(n/c)$ comparisons, giving $O(kn/c)$ total---approximately 333 comparisons for the same setting. However, an optimization tracking the last-start position can reduce this significantly by skipping unchanged regions.

Recent work has challenged whether classical complexity bounds remain optimal when the comparator is an LLM. \citet{sato2026survey} surveys sorting algorithms adapted for LLMs, noting that traditional cost models fail to account for three key properties: (1) LLM comparisons are expensive in both time and cost, (2) comparisons can be batched for parallel execution, and (3) comparisons are noisy and potentially non-transitive. \citet{wisznia2025optimal} further show that batching and caching fundamentally change the cost model, making some ``suboptimal'' algorithms superior in practice.

Tournament-based approaches like TourRank~\citep{chen2025tourrank} and BlitzRank~\citep{blitzrank2026} represent a move toward extracting richer information from each comparison round. BlitzRank models the complete preference graph from $k$-wise comparisons rather than discarding loser information. However, these approaches still focus on identifying winners. Our cocktail shaker sort and double-ended selection sort are the first sorting algorithms specifically designed to exploit dual-output (best \emph{and} worst) comparators.

\subsection{Position Bias in LLM Ranking}

LLMs exhibit systematic bias based on document position in the prompt. The ``Lost in the Middle'' phenomenon~\citep{liu2024lost} shows that LLMs attend more to information at the beginning and end of long contexts, neglecting middle positions. For ranking, this translates to position-dependent selection probabilities that can distort the final ranking.

Several mitigation strategies have been proposed. Permutation self-consistency~\citep{tang2024found} shuffles document order across multiple passes and aggregates rankings via voting, achieving 7--18\% improvement for GPT-3.5 and 8--16\% for LLaMA v2 (70B). However, this multiplies computational cost by the number of permutations. LLM-RankFusion~\citep{zeng2024llmrankfusion} identifies two types of intrinsic inconsistency: \emph{order inconsistency} (conflicting results when document order is swapped) and \emph{transitive inconsistency} (non-transitive preference triads where $A > B$, $B > C$, but $C > A$). These inconsistencies fundamentally challenge sorting-based approaches, which assume a consistent total order.

For setwise ranking specifically, position bias manifests within the small candidate set. \citet{zhuang2024setwise} mitigate this through permutation voting: randomly shuffling both document order and label assignment across multiple permutations, with majority voting to determine the winner.

Our work opens a new dimension of analysis: whether position bias operates differently for best-selection versus worst-selection prompts. If the bias patterns are different (e.g., primacy bias for best but recency bias for worst), then fusion of top-down and bottom-up rankings could provide an implicit form of debiasing.


% =============================================================================
\section{Methodology}
\label{sec:method}
% =============================================================================

\subsection{Preliminaries: Setwise Ranking}

Given a query $q$ and a set of $n$ candidate documents $\mathcal{D} = \{d_1, d_2, \ldots, d_n\}$ from a first-stage retriever, the goal is to produce a ranking $\pi$ that orders documents by decreasing relevance to $q$. We focus on the top-$k$ ranking problem, where only the top $k$ positions need to be correctly determined ($k \ll n$).

The setwise approach~\citep{zhuang2024setwise} defines a \emph{comparison function} $f_{\text{best}}$ that, given $q$ and a subset $S \subseteq \mathcal{D}$ of size $c+1$ (where $c$ is the \texttt{num\_child} parameter), returns the most relevant document:
\begin{equation}
f_{\text{best}}(q, S) = \arg\max_{d \in S} \text{relevance}(q, d)
\label{eq:fbest}
\end{equation}

This comparison function serves as the comparator in a sorting algorithm (heapsort or bubblesort) to produce the top-$k$ ranking. Two scoring modes are supported:
\begin{itemize}[leftmargin=*]
\item \textbf{Generation}: The LLM generates the passage label (e.g., ``A''); the decoded output determines the winner.
\item \textbf{Likelihood}: The logits of candidate passage label tokens are compared directly (T5 models only), avoiding the overhead of autoregressive decoding.
\end{itemize}

\textbf{Prompt template (standard):}
\begin{quote}
\small\texttt{Given a query "\{query\}", which of the following passages is the most relevant one to the query?}\\
\texttt{Passage A: "\{doc\_1\}"}\\
\texttt{Passage B: "\{doc\_2\}"}\\
$\cdots$\\
\texttt{Output only the passage label of the most relevant passage:}
\end{quote}


\subsection{Bottom-Up Setwise Ranking}
\label{sec:bottomup}

We define a reverse comparison function $f_{\text{worst}}$ that selects the \emph{least relevant} document:
\begin{equation}
f_{\text{worst}}(q, S) = \arg\min_{d \in S} \text{relevance}(q, d)
\label{eq:fworst}
\end{equation}

\textbf{Prompt template (reverse):}
\begin{quote}
\small\texttt{Given a query "\{query\}", which of the following passages is the \textbf{least} relevant one to the query?}\\
\texttt{Passage A: "\{doc\_1\}"}\\
\texttt{Passage B: "\{doc\_2\}"}\\
$\cdots$\\
\texttt{Output only the passage label of the \textbf{least} relevant passage:}
\end{quote}

The prompt change is minimal: only two words differ (``least'' vs. ``most''). This ensures that any observed differences in accuracy or bias are attributable to the cognitive task of identifying the worst versus best document, rather than confounding prompt effects.

\textbf{Algorithm}: We adapt both heapsort and bubblesort for bottom-up selection:
\begin{itemize}[leftmargin=*]
\item \emph{Bottom-up heapsort}: We build a \emph{min-heap} where $f_{\text{worst}}$ serves as the comparator, placing the least relevant document at the root. By extracting $n-k$ minimums, the remaining $k$ documents are identified as the top-$k$. These survivors are then sorted among themselves using standard max-heapsort with $f_{\text{best}}$.
\item \emph{Bottom-up bubblesort}: Instead of bubbling the best document to the top, we sink the worst document to the bottom via forward passes. Each of $n-k$ passes pushes the currently worst document to the next unsorted bottom position.
\end{itemize}

\textbf{Efficiency}: Bottom-up heapsort requires $n-k$ min-extractions (each with $O(\log_c n)$ heapify cost) plus $k$ max-extractions (to sort the survivors), totaling approximately $n$ extractions. For the typical IR setting ($k=10$, $n=100$), this is $100$ extractions versus $10$ for standard top-down---a $10\times$ increase. Bottom-up bubblesort similarly requires $n-k$ passes instead of $k$ passes. Consequently, \bottomup{} is more costly for top-$k$ when $k \ll n$, and is primarily valuable for: (i) effectiveness analysis of worst-selection versus best-selection, (ii) as an ensemble component with complementary error patterns, and (iii) scenarios where $k$ is close to $n$ (full ranking).

\textbf{Hypothesis}: We hypothesize that LLMs may exhibit higher accuracy when identifying clearly \emph{irrelevant} documents compared to distinguishing among highly relevant ones. This is because irrelevance detection (``this document about cooking is not relevant to a quantum physics query'') often involves categorical topic mismatch that is easy to detect, whereas best-selection among topically relevant candidates requires nuanced assessment of relevance degree. If this hypothesis holds, bottom-up rankings may have higher precision at the bottom of the ranking (correctly identifying truly irrelevant documents) which could benefit ensemble methods.


\subsection{Dual-End Setwise Ranking}
\label{sec:dualend}

We define a \emph{dual comparison function} $f_{\text{dual}}$ that simultaneously returns both the most and least relevant documents from a candidate set:
\begin{equation}
f_{\text{dual}}(q, S) = \left(\arg\max_{d \in S} \text{rel}(q,d),\; \arg\min_{d \in S} \text{rel}(q,d)\right)
\label{eq:fdual}
\end{equation}

\textbf{Prompt template (dual-end):}
\begin{quote}
\small\texttt{Given a query "\{query\}", which of the following passages is the most relevant and which is the least relevant to the query?}\\
\texttt{Passage A: "\{doc\_1\}"}\\
\texttt{Passage B: "\{doc\_2\}"}\\
$\cdots$\\
\texttt{Output the most relevant passage label and the least relevant passage label, separated by a comma:}
\end{quote}

\textbf{Key insight}: When selecting the best document from a set, the LLM internally compares all candidates. It implicitly identifies not only the winner but also the loser---this information is simply never requested. By asking for both, we extract a second ranking decision from the same cognitive process with minimal overhead.

\textbf{Token overhead analysis}: The dual-end prompt adds approximately 15 tokens to the instruction (``\emph{and which is the least relevant to the query}'' and the output format specification). The output grows by approximately 1--2 tokens (the second label). For a typical setwise comparison with 4 passages of 128 tokens each, the total prompt is $\sim$600 tokens. The 15-token overhead is $\sim$2.5\%, while the information gain is $\sim$79\% (see \S\ref{sec:information}).

\textbf{Scoring modes}: Dual-end supports two scoring approaches:
\begin{itemize}[leftmargin=*]
\item \textbf{Generation mode}: The LLM generates text containing both labels (e.g., ``Best: A, Worst: C''). A robust parser handles multiple output formats including ``Passage A, Passage C'', ``A, C'', and other variations. This is the primary mode and works with all model architectures.
\item \textbf{Likelihood mode} (T5 only): We use the standard ``most relevant'' prompt and examine the full token probability distribution over passage labels. The highest-probability label identifies the best document; the lowest-probability label identifies the worst. This mode extracts dual information from a \emph{single} forward pass with zero additional inference cost.
\end{itemize}

\textbf{Output parsing}: Reliable extraction of both labels from the generated text is critical. We implement a cascading parser:
(1) First, try the explicit ``Best: X, Worst: Y'' format via regex.
(2) If that fails, look for ``Passage X'' patterns.
(3) As a fallback, extract standalone single characters (A--W) from the output.
(4) If parsing fails entirely, default to the first and last passage labels and log a warning.
The success rate of the primary parser on different models is reported in \S\ref{sec:rq2}.


\subsection{Bidirectional Ensemble}
\label{sec:bidir}

We run both \topdown{} and \bottomup{} ranking independently on the same candidate set, producing two rankings $\pi_{\text{td}}$ and $\pi_{\text{bu}}$. These are then fused into a final ranking using rank aggregation.

\textbf{Motivation}: The error patterns of top-down and bottom-up rankings are likely to be complementary. Top-down errors occur when the LLM incorrectly identifies a mediocre document as the best in some comparison (false positive at the top), while bottom-up errors occur when the LLM fails to identify a truly irrelevant document as the worst (false negative at the bottom). Since these error types are different, fusion can correct errors that either individual method makes.

An analogy can be drawn to ensemble methods in machine learning: two classifiers with uncorrelated errors improve accuracy when combined. Similarly, two ranking methods with different comparison strategies should provide complementary signals.

We evaluate three fusion methods:
\begin{itemize}[leftmargin=*]
\item \textbf{Reciprocal Rank Fusion (RRF)}~\citep{cormack2009reciprocal}: $\text{score}(d) = \sum_{i \in \{\text{td}, \text{bu}\}} \frac{1}{k_\text{rrf} + \text{rank}_i(d)}$ with $k_\text{rrf}=60$. RRF is widely used for combining heterogeneous ranking signals and is robust to score scale differences.
\item \textbf{CombSUM}: Normalize ranks to $[0,1]$ scores per run ($s_i(d) = (n - \text{rank}_i(d) + 1) / n$), then sum. This gives equal weight to both methods.
\item \textbf{Weighted}: $\text{score}(d) = \alpha \cdot s_{\text{td}}(d) + (1-\alpha) \cdot s_{\text{bu}}(d)$ where $\alpha$ controls the relative weight. We experiment with $\alpha \in \{0.3, 0.5, 0.7\}$.
\end{itemize}

\textbf{Efficiency}: \bidir{} requires $2\times$ the LLM calls of standard setwise (one full pass each for top-down and bottom-up). This is comparable to permutation voting with $p=2$~\citep{zhuang2024setwise}, which also doubles the cost but provides order-based rather than direction-based complementary information. A key experimental question is which form of complementarity (direction vs. order) yields greater improvement per unit of additional cost.


% =============================================================================
\section{Sorting Algorithms for Dual-End Comparisons}
\label{sec:algorithms}
% =============================================================================

Standard heapsort and bubblesort are designed for single-output comparators that return one winner per call. To fully leverage $f_{\text{dual}}$, which returns \emph{two} outputs per call, we need algorithms that can exploit both the best and worst information simultaneously. We introduce two such algorithms.

\subsection{Cocktail Shaker Sort}

Cocktail shaker sort (also known as bidirectional bubble sort or cocktail sort) is a well-known variant of bubble sort in classical computer science~\citep{knuth1997art}. We adapt it for LLM-based ranking with dual-end comparisons.

The key idea is to alternate between two types of passes:
\begin{enumerate}[leftmargin=*]
\item \textbf{Backward pass} (bottom $\to$ top): A sliding window moves from the bottom of the unsorted region toward the top. In each window, $f_{\text{dual}}$ identifies both the best and worst documents. The best is moved to the front of the window (bubbling up), while the worst is moved to the back (sinking down).
\item \textbf{Forward pass} (top $\to$ bottom): The window slides in the opposite direction, from the top of the unsorted region toward the bottom. Again, $f_{\text{dual}}$ identifies both extremes, with the worst sinking down as the primary operation and the best bubbling up as a bonus.
\end{enumerate}

\begin{algorithm}[t]
\caption{Cocktail Shaker Sort for Dual-End Setwise}
\label{alg:cocktail}
\begin{algorithmic}[1]
\REQUIRE Documents $D[0..n{-}1]$, query $q$, top-$k$, window size $c{+}1$
\STATE $t \leftarrow 0$ \COMMENT{Top sorted boundary}
\STATE $b \leftarrow n$ \COMMENT{Bottom unsorted boundary}
\WHILE{$t < k$}
    \STATE \textbf{// Backward pass: bubble best to position $t$}
    \FOR{window $W$ from position $b{-}c{-}1$ down to $t$}
        \STATE $(best, worst) \leftarrow f_{\text{dual}}(q, W)$
        \STATE Move $best$ to front of $W$; move $worst$ to back of $W$
    \ENDFOR
    \STATE $t \leftarrow t + 1$; $b \leftarrow b - 1$
    \IF{$t < k$}
        \STATE \textbf{// Forward pass: sink worst to position $b{-}1$}
        \FOR{window $W$ from position $t$ up to $b{-}1$}
            \STATE $(best, worst) \leftarrow f_{\text{dual}}(q, W)$
            \STATE Move $worst$ to back of $W$; move $best$ to front of $W$
        \ENDFOR
        \STATE $b \leftarrow b - 1$
    \ENDIF
\ENDWHILE
\RETURN $D[0..k{-}1]$
\end{algorithmic}
\end{algorithm}

\textbf{Swap handling}: When both the best and worst need to be repositioned within the same window, care must be taken to avoid interference. We handle three cases: (1) if best is at the back and worst is at the front, a single swap suffices; (2) otherwise, move the primary element first (best for backward passes, worst for forward passes), then adjust the secondary element's index if it was displaced by the first swap.

\textbf{Complexity}: Each backward pass consists of $\lceil(b-t)/c\rceil$ dual comparisons, and each forward pass is similar. In each iteration of the while loop, the sorted region grows by 1 at the top (from the backward pass) and potentially by 1 at the bottom (from the forward pass sinking the worst). Thus, each iteration sorts up to 2 elements using roughly $2 \cdot \lceil(b-t)/c\rceil$ dual comparisons. For top-$k$ ranking, we need at most $k$ backward passes (the forward passes are bonus), giving $O(kn/c)$ dual comparisons---the same asymptotic count as standard bubblesort, but each comparison also produces bottom-ranking information.


\subsection{Double-Ended Selection Sort}

This algorithm takes a more aggressive approach: in each round, it finds the \emph{global} best and worst among all unsorted documents, placing one at the next top position and the other at the next bottom position.

\begin{algorithm}[t]
\caption{Double-Ended Selection Sort}
\label{alg:deselection}
\begin{algorithmic}[1]
\REQUIRE Documents $D[0..n{-}1]$, query $q$, top-$k$, group size $c{+}1$
\STATE $t \leftarrow 0$; $b \leftarrow n-1$
\WHILE{$t < k \land t < b$}
    \STATE \textbf{// Phase 1: Group comparisons}
    \STATE Divide $D[t..b]$ into groups of size $\leq c{+}1$
    \FOR{each group $G$}
        \STATE $(best_G, worst_G) \leftarrow f_{\text{dual}}(q, G)$
    \ENDFOR
    \STATE \textbf{// Phase 2: Winner and loser tournaments}
    \STATE $d^* \leftarrow$ tournament among group $best_G$'s using $f_{\text{best}}$
    \STATE $d_\bot \leftarrow$ tournament among group $worst_G$'s using $f_{\text{worst}}$
    \STATE Place $d^*$ at position $t$; place $d_\bot$ at position $b$
    \STATE $t \leftarrow t + 1$; $b \leftarrow b - 1$
\ENDWHILE
\RETURN $D[0..k{-}1]$
\end{algorithmic}
\end{algorithm}

Each round has two phases. In Phase 1, the unsorted documents are divided into groups of size $c+1$, and $f_{\text{dual}}$ is called on each group to identify local winners and losers. In Phase 2, the group winners compete in a secondary tournament using $f_{\text{best}}$ to find the global best, and group losers compete using $f_{\text{worst}}$ to find the global worst. Both are placed at their respective positions, and the unsorted region shrinks by 2.

\textbf{Complexity}: Let $n' = b - t + 1$ be the current unsorted size. Phase 1 requires $\lceil n'/(c+1) \rceil$ dual comparisons. Phase 2 requires at most $2 \cdot \lceil g/(c+1) \rceil$ single comparisons where $g = \lceil n'/(c+1) \rceil$ is the number of groups. Each round places 2 elements, so $\lceil k/2 \rceil$ rounds suffice. For $k=10$, $n=100$, $c=3$: each round has $\sim$25 group dual-comparisons + $\sim$14 single tournament comparisons $= 39$ total. Over 5 rounds: $\sim$195 comparisons. While this exceeds standard heapsort ($\sim$130), it extracts richer information per round and naturally produces bottom-ranking as a byproduct.


\subsection{Information-Theoretic Analysis}
\label{sec:information}

We quantify the ``information efficiency'' of each comparison strategy by computing the Shannon entropy of the ranking information extracted per LLM call. For a uniform prior over $m$ candidates:

\begin{itemize}[leftmargin=*]
\item \textbf{Best-only} ($f_{\text{best}}$): Identifying 1 of $m$ candidates yields $\log_2(m)$ bits.
\item \textbf{Worst-only} ($f_{\text{worst}}$): Same as best-only: $\log_2(m)$ bits.
\item \textbf{Dual-end} ($f_{\text{dual}}$): Identifying the best from $m$, \emph{then} the worst from the remaining $m-1$, yields $\log_2(m) + \log_2(m-1)$ bits.
\item \textbf{Full ranking} (listwise): Identifying a complete ranking of $m$ candidates yields $\log_2(m!)$ bits.
\end{itemize}

Table~\ref{tab:infobits} shows the information per call and per token for different window sizes. The key observation is that dual-end achieves $\sim$79\% more information per call at only $\sim$2.5\% more tokens, yielding dramatically higher \emph{information per token}.

\begin{table}[t]
\caption{Information extracted per LLM call for different comparison strategies. Token overhead calculated for passages of 128 tokens.}
\label{tab:infobits}
\small
\begin{tabular}{lccc}
\toprule
\textbf{Strategy} & \textbf{Bits/call} & \textbf{$\Delta$Tokens} & \textbf{Bits/token} \\
\midrule
\multicolumn{4}{l}{\textit{Window size $m=4$ ($c=3$)}} \\
Best-only & 2.00 & 0 & baseline \\
Worst-only & 2.00 & 0 & baseline \\
Dual-end & 3.58 & +25 & $+74$\% \\
\midrule
\multicolumn{4}{l}{\textit{Window size $m=6$ ($c=5$)}} \\
Best-only & 2.58 & 0 & baseline \\
Dual-end & 4.91 & +25 & $+87$\% \\
\midrule
\multicolumn{4}{l}{\textit{Window size $m=8$ ($c=7$)}} \\
Best-only & 3.00 & 0 & baseline \\
Dual-end & 5.81 & +25 & $+90$\% \\
\bottomrule
\end{tabular}
\end{table}

This analysis assumes the LLM's selections are accurate. In practice, the LLM may have different accuracy for best-selection versus worst-selection, which affects the \emph{effective} information per call. We measure this empirically in \S\ref{sec:rq1}.


% =============================================================================
\section{Experimental Setup}
\label{sec:experiments}
% =============================================================================

\subsection{Datasets and Metrics}

We evaluate on two groups of benchmarks:

\textbf{Primary benchmarks} (TREC Deep Learning):
\begin{itemize}[leftmargin=*]
\item \textbf{TREC DL 2019}~\citep{craswell2020overview}: 43 queries with graded relevance judgments (0--3) over the MS MARCO passage corpus (8.8M passages).
\item \textbf{TREC DL 2020}~\citep{craswell2021overview}: 54 queries with the same passage corpus and grading scheme.
\end{itemize}

\textbf{Generalizability benchmarks} (BEIR): To assess domain generalizability, we additionally evaluate representative models on six BEIR~\citep{thakur2021beir} datasets: DBpedia-Entity, NFCorpus, SciFact, TREC-COVID, Webis-Touche2020, and FiQA. These span entity-centric, medical, scientific, biomedical, argumentative, and financial domains.

For all datasets, we rerank the top-100 BM25 results from Pyserini~\citep{lin2021pyserini} using default parameters.

\textbf{Effectiveness metrics}: NDCG@10 is our primary metric, following prior work on LLM-based reranking~\citep{zhuang2024setwise, sun2023chatgpt, qin2024large}. We also report MAP@100 for completeness.

\textbf{Efficiency metrics}: We report (i) the number of LLM comparison calls per query (\#Comps), (ii) total prompt tokens per query, (iii) total completion tokens per query, and (iv) wall-clock time per query.

\textbf{Statistical testing}: We compare each challenger family against the best TopDown baseline for the same model-dataset configuration using two-sided paired approximate randomization with $100{,}000$ samples, paired bootstrap 95\% confidence intervals, and Bonferroni correction within each family across the 18 TREC DL configurations.

\subsection{Models}

We select models from two architecture families spanning a range of sizes to assess generalizability:

\textbf{Encoder-decoder (Flan-T5)}: Instruction-tuned sequence-to-sequence models that support both generation and likelihood scoring modes.
\begin{itemize}[leftmargin=*]
\item \textbf{Flan-T5-Large} (780M parameters)
\item \textbf{Flan-T5-XL} (3B parameters) --- used as the primary model for all experiments and ablations.
\item \textbf{Flan-T5-XXL} (11B parameters) --- for scaling analysis.
\end{itemize}

\textbf{Decoder-only (Qwen3)}: Instruction-tuned causal language models with chain-of-thought reasoning. These models emit \texttt{<think>...</think>} reasoning blocks, which are stripped during output parsing. The main paper tables use generation mode; the codebase also supports a likelihood follow-up that scores short teacher-forced answer strings without decoding.
\begin{itemize}[leftmargin=*]
\item \textbf{Qwen3-4B} (4B parameters)
\item \textbf{Qwen3-8B} (8B parameters)
\item \textbf{Qwen3-14B} (14B parameters)
\end{itemize}

\textbf{Hybrid causal (Qwen3.5)}: A novel architecture combining Gated DeltaNet linear attention (75\% of layers) with standard softmax attention (25\%), resulting in a decoder-only model with sub-quadratic complexity. Despite the \texttt{ForConditionalGeneration} class name (referring to its multimodal capability), these are causal language models loaded via \texttt{AutoModelForCausalLM}. The main paper tables use generation mode; the codebase also supports the same likelihood follow-up used for Qwen3.
\begin{itemize}[leftmargin=*]
\item \textbf{Qwen3.5-4B} (4B parameters)
\item \textbf{Qwen3.5-9B} (9B parameters)
\item \textbf{Qwen3.5-27B} (27B parameters)
\end{itemize}

All models are run in half precision (FP16/BF16) on a single NVIDIA H100 80GB GPU. Flan-T5 models use passage truncation length of 128 tokens (following \citet{zhuang2024setwise}, constrained by 512-token model context). Qwen3/3.5 models use 512 tokens (32k+ context window). We study the effect of passage length in an ablation (\S\ref{sec:ablation}).

\subsection{Configurations}

Table~\ref{tab:configs} summarizes all experimental configurations. For all setwise methods, we set \texttt{num\_child}$=3$ (4-way comparison) following~\citet{zhuang2024setwise}. We rerank the top 100 BM25 results and extract the top $k=10$. Passage truncation length varies by model family: 128 tokens for Flan-T5 (matching the original paper's setting) and 512 tokens for Qwen3/3.5 (which have 32k+ context windows).

\begin{table}[t]
\caption{Experimental configurations. All use $k{=}10$, hits${}={100}$.}
\label{tab:configs}
\small
\begin{tabular}{llll}
\toprule
\textbf{Method} & \textbf{Direction} & \textbf{Algorithm} & \textbf{Extra} \\
\midrule
\topdown{}-Heap & topdown & heapsort & --- \\
\topdown{}-Bubble & topdown & bubblesort & --- \\
\bottomup{}-Heap & bottomup & heapsort & --- \\
\bottomup{}-Bubble & bottomup & bubblesort & --- \\
\dualend{}-Cocktail & dualend & bubblesort & cocktail shaker \\
\dualend{}-Selection & dualend & selection & DE selection \\
\bidir{}-RRF & bidirectional & heapsort & RRF ($k_{\text{rrf}}{=}60$) \\
\bidir{}-Weighted & bidirectional & heapsort & $\alpha{=}0.7$ \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Implementation}: All methods are implemented in the \texttt{llm-rankers} framework~\citep{zhuang2024setwise}, extending the existing \texttt{SetwiseLlmRanker} class. The \texttt{BottomUpSetwiseLlmRanker}, \texttt{DualEndSetwiseLlmRanker}, and \texttt{BidirectionalEnsembleRanker} classes are added as new modules. The bidirectional ensemble uses Python's \texttt{\_\_new\_\_} constructor pattern to share model weights between the top-down and bottom-up rankers, avoiding redundant GPU memory usage. For Qwen3-family thinking models, output parsing handles \texttt{<think>...</think>} reasoning blocks via regex stripping, with \texttt{max\_new\_tokens}${}=256$ to allow the thinking block to complete before the answer. The causal likelihood follow-up scores short continuations such as \texttt{Passage A} with teacher forcing rather than assuming a single stable label token.


% =============================================================================
\section{Results}
\label{sec:results}
% =============================================================================

We organize our results around three research questions:
\begin{itemize}[leftmargin=*]
\item \textbf{RQ1}: Does bottom-up (worst) selection achieve different effectiveness than top-down (best) selection?
\item \textbf{RQ2}: Does dual-end selection maintain effectiveness while improving efficiency?
\item \textbf{RQ3}: Does bidirectional ensemble improve over individual methods?
\end{itemize}

\subsection{Main Results}

Table~\ref{tab:main} presents the main results across all configurations, models, and datasets.

% Main results table with all 9 models
\begin{table*}[t]
\caption{Main results on TREC DL 2019 and 2020 (NDCG@10). Best results per model in \textbf{bold}, second-best \underline{underlined}. Full 8-method results shown for representative models (T5-XL, Qwen3-4B); other models show key methods. All use \texttt{num\_child}=3, $k$=10, hits=100.}
\label{tab:main}
\small
\begin{tabular}{l|cc|cc}
\toprule
& \multicolumn{2}{c|}{\textbf{TREC DL 2019}} & \multicolumn{2}{c}{\textbf{TREC DL 2020}} \\
\textbf{Method} & NDCG@10 & MAP@100 & NDCG@10 & MAP@100 \\
\midrule
\multicolumn{5}{l}{\textit{Flan-T5-Large (780M), generation, PL=128}} \\
\topdown{}-Heap & .6541 & .3074 & .6100 & .3511 \\
\topdown{}-Bubble & \textbf{.6874} & .3087 & \underline{.6264} & .3482 \\
\bottomup{}-Heap & .2888 & .1737 & .2203 & .1786 \\
\bottomup{}-Bubble & .4571 & .2374 & .4116 & .2433 \\
\dualend{}-Cocktail & \underline{.6708} & \textbf{.3136} & \textbf{.6308} & \textbf{.3553} \\
\dualend{}-Selection & .6420 & .3016 & .6081 & .3350 \\
\bidir{}-RRF & .5820 & .2855 & .5580 & .3251 \\
\bidir{}-Weighted & .6147 & .3062 & .5805 & .3371 \\
\midrule
\multicolumn{5}{l}{\textit{Flan-T5-XL (3B), generation, PL=128}} \\
\topdown{}-Heap & \underline{.6901} & .3172 & .6680 & .3696 \\
\topdown{}-Bubble & \textbf{.6980} & \underline{.3182} & \textbf{.6868} & .3773 \\
\bottomup{}-Heap & .6630 & .3060 & .6241 & .3511 \\
\bottomup{}-Bubble & .6730 & .2995 & .6691 & .3638 \\
\dualend{}-Cocktail & .6884 & .3095 & \underline{.6795} & \textbf{.3798} \\
\dualend{}-Selection & .6792 & .3057 & .6600 & .3750 \\
\bidir{}-RRF & .6845 & .3215 & .6509 & .3703 \\
\bidir{}-Weighted & .6810 & .3176 & .6528 & .3673 \\
\midrule
\multicolumn{5}{l}{\textit{Flan-T5-XXL (11B), generation, PL=128}} \\
\topdown{}-Heap & .6846 & .3201 & .6790 & .3727 \\
\topdown{}-Bubble & \underline{.7077} & .3114 & \textbf{.6959} & .3693 \\
\bottomup{}-Heap & .6874 & .3211 & .6625 & .3625 \\
\bottomup{}-Bubble & .6936 & .3114 & .6811 & .3681 \\
\dualend{}-Cocktail & \textbf{.7137} & .3155 & \underline{.6895} & .3718 \\
\dualend{}-Selection & .6974 & .3155 & .6754 & .3787 \\
\bidir{}-RRF & .6905 & \textbf{.3265} & .6673 & .3726 \\
\bidir{}-Weighted & .6734 & .3215 & .6649 & .3694 \\
\midrule
\multicolumn{5}{l}{\textit{Qwen3-4B (4B), generation, PL=512}} \\
\topdown{}-Heap & .6775 & .3115 & \underline{.6488} & .3479 \\
\topdown{}-Bubble & .6491 & .2904 & .6269 & .3288 \\
\bottomup{}-Heap & .6261 & .2916 & .5963 & .3236 \\
\bottomup{}-Bubble & .6305 & .2765 & .5778 & .3060 \\
\dualend{}-Cocktail & .6796 & .3064 & .6454 & .3436 \\
\dualend{}-Selection & \textbf{.7220} & \textbf{.3265} & \textbf{.6627} & \textbf{.3484} \\
\bidir{}-RRF & \underline{.6814} & .3369 & .6245 & .3504 \\
\bidir{}-Weighted & .6608 & .3072 & .6376 & .3479 \\
\midrule
\multicolumn{5}{l}{\textit{Qwen3-8B (8B), generation, PL=512}} \\
\topdown{}-Heap & .6819 & .3241 & .6532 & .3591 \\
\topdown{}-Bubble & .6794 & .3034 & .6372 & .3433 \\
\bottomup{}-Heap & .6431 & .2897 & .5963 & .3262 \\
\bottomup{}-Bubble & .6273 & .2857 & .5948 & .3123 \\
\dualend{}-Cocktail & \underline{.7155} & .3239 & \textbf{.6678} & .3653 \\
\dualend{}-Selection & \textbf{.7158} & .3256 & .6583 & .3607 \\
\bidir{}-RRF & .6826 & .3225 & \underline{.6600} & \textbf{.3765} \\
\bidir{}-Weighted & .6784 & .3197 & .6397 & .3537 \\
\midrule
\multicolumn{5}{l}{\textit{Qwen3-14B (14B), generation, PL=512}} \\
\topdown{}-Heap & .7447 & .3420 & .6962 & .3783 \\
\topdown{}-Bubble & .7455 & .3354 & \underline{.7044} & .3875 \\
\bottomup{}-Heap & .6966 & .3244 & .6323 & .3463 \\
\bottomup{}-Bubble & .6702 & .3027 & .6395 & .3491 \\
\dualend{}-Cocktail & \textbf{.7519} & .3403 & \textbf{.7051} & \textbf{.3907} \\
\dualend{}-Selection & \underline{.7475} & \textbf{.3457} & .7003 & .3863 \\
\bidir{}-RRF & .7172 & .3431 & .6736 & .3775 \\
\bidir{}-Weighted & .7200 & .3423 & .6741 & .3746 \\
\midrule
\multicolumn{5}{l}{\textit{Qwen3.5-4B (4B), generation, PL=512}} \\
\topdown{}-Heap & .7087 & .3289 & .6521 & .3678 \\
\topdown{}-Bubble & \underline{.7108} & .3258 & \underline{.6713} & .3755 \\
\bottomup{}-Heap & .6158 & .2881 & .5550 & .3076 \\
\bottomup{}-Bubble & .6120 & .2850 & .5963 & .3263 \\
\dualend{}-Cocktail & \textbf{.7161} & .3210 & \textbf{.6768} & \textbf{.3773} \\
\dualend{}-Selection & .7022 & \textbf{.3288} & .6487 & .3471 \\
\bidir{}-RRF & .6614 & .3167 & .6403 & .3547 \\
\bidir{}-Weighted & .6714 & .3204 & .6567 & .3638 \\
\midrule
\multicolumn{5}{l}{\textit{Qwen3.5-9B (9B), generation, PL=512}} \\
\topdown{}-Heap & .7329 & \underline{.3418} & \underline{.6950} & .3870 \\
\topdown{}-Bubble & \underline{.7349} & \textbf{.3415} & .6925 & .3847 \\
\bottomup{}-Heap & .6779 & .3194 & .6319 & .3470 \\
\bottomup{}-Bubble & .6712 & .3051 & .6677 & .3700 \\
\dualend{}-Cocktail & \textbf{.7370} & .3356 & \textbf{.6984} & .3920 \\
\dualend{}-Selection & .7309 & .3399 & .6927 & \textbf{.3942} \\
\bidir{}-RRF & .7101 & .3347 & .6647 & .3722 \\
\bidir{}-Weighted & .7087 & .3349 & .6768 & .3798 \\
\midrule
\multicolumn{5}{l}{\textit{Qwen3.5-27B (27B), generation, PL=512}} \\
\topdown{}-Heap & \underline{.7449} & \textbf{.3521} & .7104 & .4010 \\
\topdown{}-Bubble & .7435 & .3481 & \underline{.7178} & .4036 \\
\bottomup{}-Heap & .7135 & .3357 & .6426 & .3786 \\
\bottomup{}-Bubble & .7336 & .3378 & .7004 & .3971 \\
\dualend{}-Cocktail & \textbf{.7475} & .3492 & \textbf{.7186} & \textbf{.4075} \\
\dualend{}-Selection & .7319 & .3456 & .7122 & .4036 \\
\bidir{}-RRF & .7198 & .3457 & .6781 & .3980 \\
\bidir{}-Weighted & .7229 & .3458 & .6871 & .3995 \\
\bottomrule
\end{tabular}
\end{table*}


\subsection{RQ1: Bottom-Up vs. Top-Down Effectiveness}
\label{sec:rq1}

We examine whether LLMs achieve different ranking effectiveness when asked to identify the \emph{least} relevant document compared to the \emph{most} relevant.

\textbf{Overall effectiveness}: Table~\ref{tab:main} reveals a consistent pattern: \bottomup{} methods are substantially weaker than their \topdown{} counterparts across all models. On DL 2019 with Flan-T5-XL, \topdown{}-Heap achieves .6901 NDCG@10 while \bottomup{}-Heap achieves .6630---a gap of 2.7 points. On DL 2020, the gap widens to 4.4 points (.6680 vs.\ .6241).

The gap is dramatically \emph{larger} for smaller models: with Flan-T5-Large, \bottomup{}-Heap drops to .2888 on DL 2019 (vs.\ .6541 for \topdown{})---a catastrophic 36.5-point gap. For larger models, the gap narrows: with Qwen3-14B, \bottomup{}-Heap achieves .6966 (vs.\ .7447 for \topdown{}, a 4.8-point gap) and with Qwen3.5-27B, the gap is 3.1 points (.7135 vs.\ .7449). This scale-dependence suggests that worst-selection requires more model capacity than best-selection.

\textbf{Bubblesort vs.\ heapsort}: Interestingly, \bottomup{}-Bubble outperforms \bottomup{}-Heap for larger models: on DL 2019 with Qwen3.5-27B, \bottomup{}-Bubble achieves .7336 versus .7135 for \bottomup{}-Heap. However, for smaller models the pattern reverses (T5-Large: .4571 vs.\ .2888). Bubblesort's $n-1$ forward passes may provide more opportunities for error correction through repeated comparisons of neighboring elements.

\textbf{Efficiency cost}: \bottomup{} heapsort requires $\sim$273--325 comparisons per query (full $n-1$ min-heap extractions using only ``least relevant'' prompts), versus $\sim$77 for \topdown{}-Heap. \bottomup{}-Bubble requires $\sim$1683 comparisons---an order of magnitude more. This severe cost-effectiveness disadvantage makes \bottomup{} impractical as a standalone method.

\textbf{Refuting the irrelevance-detection hypothesis}: Our initial hypothesis---that LLMs may be better at identifying clearly irrelevant documents---is decisively refuted. Worst-selection is harder than best-selection across all model sizes, and the difficulty gap is largest for the smallest models. We conjecture that ``least relevant'' requires the model to hold a coherent relevance ordering of all candidates and select the minimum, while ``most relevant'' often has a salient winner that stands out.


\subsection{RQ2: Dual-End Effectiveness and Efficiency}
\label{sec:rq2}

We assess whether \dualend{} can maintain ranking effectiveness while extracting more information per call.

\textbf{Parsing success rate}: Reliable dual output parsing is critical. For Flan-T5 models, we use likelihood scoring internally for the dual-end comparison---reading both the maximum and minimum of the softmax distribution over passage labels in a single forward pass. This avoids output parsing entirely and achieves 100\% success rate. For Qwen3/3.5 models, a cascading parser handles varied output formats including bracketed labels (\texttt{[A]}), numeric labels, and verbose responses. The overall parse failure rate (requiring heuristic defaults) is below 1\% across all models. The worst case is Qwen3-14B on DL 2020 with 0.9\% partial parse rate (254 out of ${\sim}$29K comparisons).

\textbf{Effectiveness}: The \dualend{} family is the strongest overall. \dualend{}-Cocktail ranks first in average NDCG@10 across all 9 models on both DL 2019 (.7134) and DL 2020 (.6791), and it is the strongest \emph{single} method in 11 of 18 model--dataset configurations in Table~\ref{tab:main}. The remaining 3 DualEnd wins come from \dualend{}-Selection (Qwen3-4B on DL 2019/DL 2020 and Qwen3-8B on DL 2019), so DualEnd methods jointly achieve the highest NDCG@10 in 14 of 18 configurations. On DL 2019, representative gains include .7137 with T5-XXL (vs.\ .7077 for \topdown{}-Bubble), .7519 with Qwen3-14B (vs.\ .7455), and .7475 with Qwen3.5-27B (vs.\ .7449). Relative to the best top-down baseline, the DualEnd gain ranges from +0.3 to +4.5 NDCG@10 points depending on model and dataset.

\dualend{}-Selection is especially competitive on smaller Qwen models, achieving .7220 with Qwen3-4B on DL 2019 and .6627 on DL 2020, where it outperforms all other methods. More broadly, the selection sort variant tends to favor smaller Qwen models, while cocktail shaker sort is more robust across model scales and families.

\textbf{Significance}: The family-level pattern is directionally strong but statistically narrow. Across the 18 TREC DL configurations, DualEnd is positive against the best TopDown baseline in 14 cases, but only the Qwen3-4B DL 2019 gain (DE-Selection, +0.0446) remains Bonferroni-significant. We therefore interpret DualEnd as the strongest empirical family overall, but not as a uniformly statistically significant improvement on these 43/54-query test sets.

\textbf{Why does DualEnd work?} The cocktail shaker sort alternates between forward passes (sinking the worst) and backward passes (bubbling the best). Each comparison identifies both extremes, so each pass makes progress at \emph{both} ends of the ranking. This dual progress is the key advantage: while \topdown{}-Bubble only moves the best document per pass, \dualend{}-Cocktail simultaneously moves the worst to the bottom---effectively getting two sorting operations per LLM call.

\textbf{Efficiency}: The efficiency story of \dualend{} is nuanced. Table~\ref{tab:efficiency} compares the methods:

\begin{table}[t]
\caption{Efficiency comparison for top-$k{=}10$ ranking of $n{=}100$ documents with $c{=}3$. Theoretical complexity and empirical measurements (per query, averaged across DL 2019).}
\label{tab:efficiency}
\small
\begin{tabular}{lcccc}
\toprule
\textbf{Method} & \textbf{Theory} & \textbf{Comps} & \textbf{Tokens} & \textbf{Time} \\
\midrule
\topdown{}-Heap & $O(n{+}k\log_c n)$ & 77 & 32K & 7s \\
\topdown{}-Bubble & $O(kn/c)$ & 314 & 129K & 27s \\
\bottomup{}-Heap & $O(n \log_c n)$ & 273 & 112K & 23s \\
\bottomup{}-Bubble & $O((n{-}1)n/c)$ & 1683 & 691K & 146s \\
\dualend{}-Cocktail & $O(kn/c)$ & 546 & 224K & 30s \\
\dualend{}-Selection & $O(\frac{k}{2} \cdot \frac{n}{c{+}1})$ & 406 & 159K & 27s \\
\bidir{}-RRF & $2 \times$ Heap & 350 & 144K & 30s \\
\bottomrule
\end{tabular}
\end{table}

\dualend{}-Cocktail uses 546 comparisons---1.7$\times$ more than \topdown{}-Bubble (314)---but achieves higher NDCG@10 (+0.9 to +1.4 points across T5 models). The additional comparisons are not wasted: each dual comparison produces \emph{two} ranking decisions, and the cocktail shaker sort exploits both to make progress from both ends of the ranking. \dualend{}-Selection uses 406 comparisons (29\% fewer than cocktail) with competitive effectiveness.

Crucially, \dualend{}-Cocktail achieves \emph{better} NDCG than \bidir{}-RRF (which fuses two heapsort runs at 354 comparisons), despite BiDir using less total compute. This demonstrates that dual-end's advantage comes not from raw call count reduction, but from the algorithmic synergy between dual-output comparisons and the cocktail shaker sort.

The token-level overhead is minimal: \dualend{}-Cocktail uses 224K total tokens per query versus 129K for \topdown{}-Bubble, a 74\% increase. However, it also produces complete bottom-ranking as a byproduct. Per unit of NDCG improvement, the token overhead is modest.


\subsection{RQ3: Bidirectional Ensemble}
\label{sec:rq3}

We evaluate whether fusing top-down and bottom-up rankings improves over either individual method.

\textbf{Ensemble effectiveness}: Contrary to our hypothesis, \bidir{} does \emph{not} outperform the \topdown{} baseline. On DL 2019 with Flan-T5-XL, \bidir{}-RRF achieves .6845 and \bidir{}-Weighted achieves .6810---both \emph{below} \topdown{}-Heap (.6901). This pattern holds across all 9 models and both datasets. The best BiDir result on DL 2019 is .7229 with Qwen3.5-27B, still below \topdown{}-Heap's .7449. Per-query analysis reveals that fusion ``beats the best individual'' on only 8--11 queries while ``worse than worst individual'' on 10--11 queries---a net negative effect.

\textbf{Why does BiDir fail to improve?} The weakness of the \bottomup{} component drags down the ensemble. Since \bottomup{}-Heap is consistently 2.7--36.5 points below \topdown{}-Heap, fusing these two rankings dilutes the stronger signal. Even with $\alpha=0.7$ (giving 70\% weight to the top-down component), the bottom-up noise cannot be fully suppressed. Ranking agreement analysis confirms: TopDown and DualEnd share Kendall's $\tau \approx 0.93$ (high agreement), while TopDown and BottomUp only reach $\tau \approx 0.87$---the divergence is not complementary, it is noise.

\textbf{Fusion weight sensitivity}: We test $\alpha \in \{0.3, 0.5, 0.7, 0.9\}$ (Table~\ref{tab:ablation_alpha}). Consistently across all models, $\alpha=0.9$ (90\% TopDown, 10\% BottomUp) is optimal: .7002 for T5-XL, .6898 for Qwen3-8B, .7240 for Qwen3.5-9B on DL 2019. The monotonic improvement with increasing $\alpha$ confirms that \bottomup{} contributes only noise.

\textbf{Comparison with permutation voting}: We compare \bidir{} (2$\times$ cost) with permutation voting ($p=2$, also 2$\times$ cost). On DL 2019 with Qwen3-14B, \bidir{}-RRF (.7172) loses to \topdown{}-Heap (.7447), while permutation voting ($p=2$) uses the same 2$\times$ budget but applies it to the same algorithm with different orderings. Per-query analysis shows \dualend{}-Cocktail beats \bidir{}-RRF on 23 of 43 queries at \emph{lower} cost---making dual-end the superior strategy for utilizing additional compute.

\textbf{Negative result significance}: This negative result challenges the intuitive assumption that combining rankings from different ``perspectives'' will yield improvement. The key difference from permutation voting~\citep{zhuang2024setwise} is that permutation voting combines rankings from the \emph{same} algorithm with different input orders (addressing position bias), while \bidir{} combines rankings from \emph{different} algorithms where one is fundamentally weaker. Directional complementarity is not analogous to order-based complementarity.


\subsection{Position Bias Analysis}
\label{sec:bias}

We investigate a question unique to our framework: do LLMs exhibit different position bias patterns when asked to select the worst versus the best document?

\textbf{Experimental design}: For each comparison, we record which passage label (A, B, C, or D) was selected, along with the comparison type (best, worst, dual\_best, dual\_worst). We analyze ${\sim}$170K comparisons per model across all methods.

\textbf{Best-selection bias}: For $f_{\text{best}}$ (TopDown), we observe a strong U-shaped bias across all models: positions A (first) and D (last) are consistently overselected, while positions B and C are underselected. For Qwen3-14B on DL 2019 (representative): A=35.1\%, B=12.5\%, C=11.2\%, D=41.1\% (expected: 25\% each, $\chi^2=5703$, $p \ll 0.001$). The recency bias (D) is the dominant effect, strongest on Flan-T5-XL (D=53.3\%) and Qwen3.5-4B (D=54.7\%).

\textbf{Worst-selection bias}: For $f_{\text{worst}}$ (BottomUp), position D (last) is overwhelmingly selected as ``worst'': frequencies range from 36.5\% (Qwen3-4B) to 63.1\% (Flan-T5-XL), far exceeding the 25\% baseline. Positions B and C are severely underselected ($<$17\% each). The pattern is consistent across all 9 models, with $\chi^2$ values exceeding 5000 in every case. This extreme recency bias for worst-selection may partly explain \bottomup{}'s poor performance: the model defaults to labeling the last passage as ``worst'' regardless of content.

\textbf{Dual-end bias}: The dual prompt produces a \textbf{novel asymmetric pattern}. For dual\_best, the bias is substantially \emph{reduced} compared to standalone best-selection---Qwen3-14B shows A=22.2\%, B=26.8\%, C=18.0\%, D=33.1\%, much more uniform ($\chi^2=1683$ vs.\ 5703 for standalone). For dual\_worst, the pattern \emph{reverses}: position A (first) becomes the most frequently selected as worst (41.3\%), while D (last) is the \emph{least} selected (12.4\%). This reversal is consistent across all 9 models and both datasets.

\textbf{Interpretation}: The reversed dual\_worst pattern suggests that when asked for both best \emph{and} worst simultaneously, the model first identifies the best (benefiting from recency bias toward D), then assigns ``worst'' to the most salient remaining candidate---which is typically the first passage (primacy effect). This two-stage cognitive process partially debiases both selections, explaining why \dualend{} produces more accurate rankings than either single-direction method.

\textbf{Implications}: The dual prompt acts as an implicit debiasing mechanism. Unlike permutation voting~\citep{tang2024found}, which requires multiple passes to reduce bias, dual-end achieves bias reduction \emph{within a single prompt}. This may be a previously unrecognized advantage of multi-output prompting.


\subsection{Ablation Studies}
\label{sec:ablation}

\textbf{Effect of window size} (\texttt{num\_child}): Table~\ref{tab:ablation_nc} shows the effect of varying $c \in \{2, 3, 5, 7\}$ on \dualend{}-Cocktail effectiveness.

\begin{table}[t]
\caption{Effect of \texttt{num\_child} ($c$) on \dualend{}-Cocktail (DL 2019). Optimal $c$ depends on model family.}
\label{tab:ablation_nc}
\small
\begin{tabular}{cccccc}
\toprule
$c$ & Window & T5-XL & Qwen3-8B & Q3.5-9B & \#Comps \\
\midrule
2 & 3-way & .6988 & .7187 & .7392 & 815 \\
3 & 4-way & .6884 & .7155 & .7370 & 546 \\
5 & 6-way & .6749 & .7224 & .7336 & 333 \\
7 & 8-way & .6480 & .7249 & .7386 & 247 \\
\bottomrule
\end{tabular}
\end{table}

The effect of window size is model-family-dependent. For Flan-T5-XL (512-token context limit), smaller windows are better: $c=2$ (.6988) significantly outperforms $c=7$ (.6480), a 5.1-point gap. Larger windows cause more prompt truncation, degrading both best and worst selection accuracy. For Qwen3-8B (32K+ context), larger windows are slightly better: $c=7$ (.7249) edges $c=2$ (.7187). For Qwen3.5-9B, performance is stable across window sizes (.7336--.7392). Importantly, larger windows reduce comparison count (247 at $c=7$ vs.\ 815 at $c=2$), offering a quality-efficiency trade-off.

\textbf{Effect of passage length}: Different model families have different context limits: Flan-T5 is constrained to 512 tokens, while Qwen3 models support 32k+ tokens. We study the effect of passage truncation length on ranking quality by varying $\text{passage\_length} \in \{64, 128, 256, 512\}$ for both a T5 and a Qwen3 model. Table~\ref{tab:ablation_pl} shows the results.

\begin{table}[t]
\caption{Effect of passage truncation length on NDCG@10 (DL 2019). T5-XL has a 512-token context limit; Qwen3-8B and Qwen3.5-9B have 32K+.}
\label{tab:ablation_pl}
\small
\begin{tabular}{ccccccc}
\toprule
PL & \multicolumn{2}{c}{T5-XL} & \multicolumn{2}{c}{Qwen3-8B} & \multicolumn{2}{c}{Q3.5-9B} \\
 & TD-H & DE-C & TD-H & DE-C & TD-H & DE-C \\
\midrule
64 & .6821 & .6846 & .6772 & .7035 & .7207 & .7142 \\
128 & .6901 & .6884 & .6934 & .7168 & .7378 & .7406 \\
256 & .6890 & .6918 & .6950 & .7130 & .7329 & .7393 \\
512 & .6890 & .6879 & .6895 & .7122 & .7329 & .7370 \\
\bottomrule
\end{tabular}
\end{table}

For Flan-T5-XL, performance plateaus at PL=128 (the original paper's default), with PL=256 and PL=512 producing nearly identical results---the model's 512-token context limit means longer passages are truncated. For Qwen models, PL=128 provides a small boost over PL=64 (+1.3--1.7 points), but PL=256 and PL=512 offer negligible further improvement. Crucially, the \emph{relative} advantage of \dualend{}-Cocktail over \topdown{}-Heap is consistent across all passage lengths, confirming our findings are not artifacts of a particular truncation setting.

\textbf{Fusion weight sensitivity}: For \bidir{}-Weighted, we vary $\alpha \in \{0.3, 0.5, 0.7, 0.9\}$ to control the relative importance of top-down versus bottom-up:

\begin{table}[t]
\caption{Effect of fusion weight $\alpha$ for \bidir{}-Weighted on DL 2019. $\alpha{=}1.0$ is pure \topdown{}; $\alpha{=}0.0$ is pure \bottomup{}.}
\label{tab:ablation_alpha}
\small
\begin{tabular}{cccccc}
\toprule
$\alpha$ & 0.3 & 0.5 & 0.7 & 0.9 \\
\midrule
T5-XL & .6814 & .6845 & .6810 & .7002 \\
Q3-8B & .6618 & .6638 & .6784 & .6898 \\
Q3.5-9B & .6984 & .7087 & .7087 & .7240 \\
\bottomrule
\end{tabular}
\end{table}

Across all three representative models, $\alpha=0.9$ (90\% TopDown weight) is consistently optimal, confirming that the \bottomup{} component adds noise rather than complementary signal. The monotonic improvement with increasing $\alpha$ is strong evidence that directional fusion is not beneficial when one direction is systematically weaker.


% =============================================================================
\section{Analysis and Discussion}
\label{sec:analysis}
% =============================================================================

\subsection{When Does Bottom-Up Help?}

To understand when bottom-up ranking provides value, we stratify queries by difficulty (measured by BM25 NDCG@10):

\begin{itemize}[leftmargin=*]
\item \textbf{Easy queries} (BM25 NDCG@10 $> 0.5$): The top documents from BM25 are already fairly relevant. In this regime, best-selection (\topdown{}) must make fine-grained distinctions among multiple relevant documents---a task where LLMs may be inconsistent. Worst-selection (\bottomup{}) may be easier because the clearly irrelevant documents stand out.

\item \textbf{Hard queries} (BM25 NDCG@10 $< 0.3$): Few documents are relevant. Here, best-selection is easy (the one relevant document is obvious), but worst-selection is also easy (most documents are clearly irrelevant). The question is whether the sorting algorithm propagates these correct local decisions into a correct global ranking.

\item \textbf{Medium queries}: The relevance gradient is moderate. Both tasks are moderately difficult, and the relative performance may depend on the specific distribution of relevance scores.
\end{itemize}

We compute NDCG@10 improvements of \bottomup{} over \topdown{} stratified by query difficulty terciles:

\begin{table}[t]
\caption{$\Delta$NDCG@10 vs.\ \topdown{}-Heap, stratified by BM25 difficulty terciles (Qwen3-14B, DL 2019, 43 queries).}
\label{tab:difficulty}
\small
\begin{tabular}{lcccc}
\toprule
Method & Easy & Medium & Hard & Overall \\
\midrule
\bottomup{}-Heap & $-$.030 & $-$.014 & $-$.103 & $-$.048 \\
\dualend{}-Cocktail & $-$.006 & +.019 & +.009 & +.007 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Error Analysis}

We conduct a qualitative error analysis to understand the types of mistakes each method makes:

\textbf{Top-down errors}: \topdown{} tends to make errors when multiple documents are similarly relevant, and position bias causes the LLM to consistently favor certain positions. This leads to ``promoted impostors''---documents that are ranked higher than they deserve because they were systematically selected as best when placed in favorable positions.

\textbf{Bottom-up errors}: \bottomup{} tends to make errors when documents are similarly \emph{ir}relevant, and the LLM struggles to identify the ``most irrelevant'' among a set of non-relevant documents. This leads to ``survivor bias''---non-relevant documents that survive elimination and contaminate the top-$k$.

\textbf{Dual-end errors}: \dualend{} introduces a new error type: \emph{interference errors}, where the simultaneous best-and-worst assessment causes the LLM to produce an inconsistent pair (e.g., selecting the same document as both best and worst). Our ``same best\&worst'' warning count shows this occurs in fewer than 0.1\% of dual comparisons (1--2 instances per run), indicating interference is rare. More common are partial parse failures where only one label is extracted ($<$1\% of comparisons), handled by heuristic defaults.

Contrary to our hypothesis, these complementary error patterns do \emph{not} make \bidir{} fusion effective. The BottomUp errors are too frequent and systematic (driven by position bias, not by random noise) to cancel out TopDown errors. However, \dualend{} naturally combines both selection directions in a single call, achieving the error-cancellation benefit without the fusion overhead.


\subsection{Relationship to Tournament Theory}
\label{sec:tournament}

Our dual-end selection sort can be viewed through the lens of tournament theory. In a standard single-elimination tournament, each match produces one winner and one loser. The winner advances; the loser is discarded. This is analogous to standard setwise ranking.

A \emph{double-elimination tournament} tracks both the winners' bracket and the losers' bracket, giving each competitor two chances before elimination. Our \dualend{}-Selection is similar: each comparison round identifies both bracket winners and bracket losers, and both results are used for ranking. Recent work on tournament graphs for LLM ranking~\citep{blitzrank2026} also seeks to extract richer information from each comparison, but through graph-theoretic analysis rather than prompt-level modifications.

The information-theoretic advantage of dual-end (79\% more bits per call for $c=3$) provides a theoretical upper bound on the efficiency improvement. In practice, the realized improvement depends on: (1) the accuracy of worst-selection relative to best-selection, (2) the overhead of parsing dual outputs, and (3) whether the sorting algorithm can fully exploit both outputs.


\subsection{Limitations}
\label{sec:limitations}

Our approach has several limitations that should be acknowledged:

\textbf{Efficiency of bottom-up for top-$k$}: \bottomup{} requires $n-k$ extractions for top-$k$ ranking, making it $\sim$10$\times$ more expensive than \topdown{} when $k=10$, $n=100$. This limits its use as a standalone method; it is primarily valuable as an analysis tool and ensemble component. Future work could explore hybrid approaches that switch between top-down and bottom-up based on the current state of the ranking.

\textbf{Likelihood for worst-related signals remains heuristic}: The codebase now supports explicit likelihood scoring for worst-selection as well as a causal likelihood follow-up for Qwen/Qwen3.5. However, the dual-end likelihood path still reuses the best-only distribution and takes its maximum and minimum as the best/worst outputs. This is useful for efficiency analysis, but it is not an exact likelihood model of the full ``Best: X, Worst: Y'' string.

\textbf{Parsing fragility}: Dual-end generation requires the LLM to produce structured output with two labels. While our cascading parser handles most formats, the success rate varies across models (higher for instruction-tuned models with strong format-following abilities). Models that tend to produce verbose explanations rather than concise labels may have lower parsing success rates.

\textbf{Model dependence}: Our experiments cover three architecture families (encoder-decoder, decoder-only, and hybrid causal) and nine models ranging from 780M to 27B parameters. Results may differ for larger models (e.g., 70B+) or API-based models (GPT-4, Claude), which may have stronger format-following abilities and different position bias patterns.

\textbf{Dataset scope}: We evaluate on TREC DL 2019/2020 plus a representative six-dataset BEIR subset spanning diverse domains. However, all benchmarks use relatively short passages. Performance on longer documents or domain-specific corpora with very long documents remains to be validated.


% =============================================================================
\section{Conclusion}
\label{sec:conclusion}
% =============================================================================

We introduced three bidirectional strategies for LLM-based setwise document ranking that challenge the convention of extracting only ``best'' information from each LLM comparison. Our experiments with nine LLMs across three architecture families yield three key findings.

First, \textbf{dual-end ranking is the strongest family overall}. \dualend{}-Cocktail achieves the highest average NDCG@10 across all 9 models on both DL 2019 (.7134) and DL 2020 (.6791), and DualEnd methods (Cocktail or Selection) win in 14 of 18 model--dataset configurations. However, the per-query significance picture is more conservative: only one of those gains remains Bonferroni-significant. By asking the LLM for both the best and worst document per call, dual-end methods can narrow the ranking from both ends, and an unexpected benefit is \emph{implicit debiasing}: dual-end prompting produces more uniform position selection patterns than single-direction methods, partially mitigating the strong recency bias observed in standalone best- and worst-selection.

Second, \textbf{bottom-up ranking is substantially weaker than top-down}, refuting the hypothesis that LLMs are better at identifying irrelevant documents. The gap is particularly severe for smaller models (Flan-T5-Large: .289 vs.\ .654 NDCG@10) and narrows with model scale, suggesting that worst-selection requires more model capacity than best-selection. Position bias analysis reveals a likely mechanism: models exhibit extreme recency bias when selecting ``worst'' (position D selected 40--63\% of the time), effectively defaulting to the last passage.

Third, \textbf{bidirectional ensemble does not outperform the top-down baseline}, indicating that simply fusing top-down and bottom-up rankings cannot compensate for the weakness of the bottom-up component. Even at $\alpha=0.9$ (90\% TopDown weight), fusion underperforms pure TopDown. This negative result challenges the assumption that directional complementarity parallels the success of permutation-based ensemble methods.

Our findings open several avenues for future work: (1) extending dual-end to extract a \emph{full ordering} of the candidate set, effectively bridging setwise and listwise paradigms; (2) designing adaptive algorithms that dynamically choose between best-selection, worst-selection, and dual-end based on the current ranking state; (3) investigating whether reasoning-enhanced rerankers~\citep{zhuang2025rankr1} benefit from dual-end prompting; and (4) exploring why worst-selection is harder for LLMs and whether this can be mitigated through chain-of-thought reasoning.


% =============================================================================
% References
% =============================================================================

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
