\documentclass[11pt]{article}

% Change "review" to "final" to generate the final (sometimes called camera-ready) version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[review]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

%Including images in your LaTeX document requires adding
%additional package(s)
\usepackage{graphicx}

% Additional packages for our paper
\usepackage{enumitem}
\setlist{leftmargin=10pt}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,decorations.pathreplacing}
\setlength{\emergencystretch}{2em}

% TeX Live 2026 + lineno v5.7 records ACL's two-column markers reversed in
% review mode, which pushes line numbers into the center gutter. Re-emit the
% column marker with the expected orientation so ACL's switch mode places line
% numbers on the outer page margins.
\makeatletter
\@ifpackageloaded{lineno}{%
  \AddToHook{build/column/before}{%
    \protected@write\@auxout{}{%
      \string\def\string\@LN@column{\if@firstcolumn2\else1\fi}%
    }%
  }%
}{}
\makeatother

% Custom commands shared across sections
\newcommand{\topdown}{\textsc{TopDown}}
\newcommand{\bottomup}{\textsc{BottomUp}}
\newcommand{\dualend}{\textsc{DualEnd}}
\newcommand{\bidir}{\textsc{BiDir}}
\newcommand{\cocktail}{\textsc{CocktailSort}}
\newcommand{\deselection}{\textsc{DESelection}}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{Finding the Worst Is Harder Than Finding the Best:\\ Directional Asymmetry in LLM Setwise Ranking}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a separate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{Anonymous Authors \\
  Anonymous Affiliation \\
  \texttt{anonymous@example.com}}

\begin{document}
\maketitle
\begin{abstract}
Large Language Models (LLMs) have emerged as effective zero-shot document rerankers. Among the four prompting paradigms---pointwise, pairwise, listwise, and setwise---the setwise approach offers a compelling trade-off between effectiveness and efficiency by asking the LLM to select the most relevant document from a candidate set. However, current setwise methods only extract one piece of ranking information per LLM call: the \emph{best} document. We investigate \emph{bidirectional} selection strategies that extract more information from each comparison. We propose three approaches: (1) \textbf{Bottom-Up Setwise Ranking}, which reverses the selection to identify the \emph{least} relevant document; (2) \textbf{Dual-End Setwise Ranking}, which asks the LLM to simultaneously identify both the most and least relevant documents; and (3) \textbf{Bidirectional Ensemble}, which fuses top-down and bottom-up rankings. We pair the dual-end strategy with two new sorting algorithms---cocktail shaker sort and double-ended selection sort---designed for dual-output comparators. Experiments on TREC DL 2019/2020 with nine LLMs spanning Flan-T5, Qwen3, and Qwen3.5 (780M--27B parameters) reveal a clear directional asymmetry: the \dualend{} family is the strongest overall (winning 14 of 18 model--dataset configurations), \bottomup{} alone is consistently weaker than \topdown{} (6 Bonferroni-significant losses), and \bidir{} ensemble does not outperform \topdown{} (3 Bonferroni-significant losses, 0 wins). Per-query significance tests are conservative on these 43/54-query test sets: only one DualEnd gain remains Bonferroni-significant, so we frame DualEnd as a quality-first refinement, not a free improvement. The mechanism we point to is positional: standalone worst-selection collapses heavily onto the last passage, while the same model placed in a joint best-and-worst prompt produces a \emph{reversed} primacy-heavy ``dual\_worst'' pattern. Joint elicitation therefore changes the model's behavior, not just its output channel.
\end{abstract}

\input{introduction}
\input{related_work}
\input{methodology}
\input{algorithms}
\input{experiment_setup}
\input{results}
\input{analysis}
\input{conclusion}
\input{limitation}

% Bibliography entries for the entire Anthology, followed by custom entries
%\bibliography{anthology,custom}
% Custom bibliography entries only
\bibliography{reference}

\appendix

\input{appendix}

\end{document}
