\section{Related Work}
\label{sec:related_work}

\paragraph{Setwise prompting and extensions.}
\citet{zhuang2024setwise} introduce setwise prompting: present a small
candidate window, ask the LLM to choose the most relevant passage, and
embed that choice inside heapsort or bubblesort. We build on the public
\texttt{llm-rankers} implementation~\citep{ielab_llmrankers} but replace
the fixed letter-label window with numeric labels so that the full live
pool is addressable. Recent setwise extensions improve other parts of
the pipeline: Setwise Insertion warm-starts sorting from the BM25 order
\citep{podolak2025setwiseinsertion}, while Rank-R1 trains the comparator
to reason before selecting within a larger setwise window
\citep{zhuang2025rankr1}. MaxContext is orthogonal: it asks what happens
when the window is the whole live pool.

\paragraph{Tournament-style reranking.}
TourRank~\citep{chen2025tourrank} and
BLITZRANK~\citep{blitzrank2026} also try to extract more value from each
multi-document judgement, using tournament structure, accumulated
evidence, or tournament graphs. These methods change how comparisons are
scheduled and aggregated. MaxContext instead changes the live-pool
granularity of a setwise selector. A matched-budget tournament comparison
would require sweeps over grouping, advancement, and graph-query knobs;
we leave that head-to-head to appendix smokes and future work rather than
making it load-bearing in a four-page short paper.

\paragraph{Listwise reranking.}
Listwise rerankers ask the model to output an ordered list, often over a
sliding window or local candidate block
\citep{sun2023rankgpt,ma2023zsl,zhang2025rankwithoutgpt,ren2025selfcalibrated}.
Their cost-quality tradeoff is therefore different from MaxContext's:
they can produce a local permutation with fewer calls, but require longer
structured outputs and face different parsing and order-sensitivity
failure modes. Our half-call claim is only against whole-pool
single-extreme setwise controls, not against listwise rerankers.

\paragraph{Long-context robustness and efficiency.}
Long contexts do not remove positional effects:
\citet{liu2024lostmiddle} show that models can under-use middle context,
and \citet{tang2024foundmiddle} improve listwise ranking with permutation
self-consistency at the cost of repeated calls. We therefore measure
input-order stability rather than assuming whole-pool prompts are
order-invariant. Finally, \citet{peng2025flopsreranking} argue that
reranker efficiency should not be summarized by proxy metrics alone. We
report calls, prompt/completion/total tokens, parser recovery, BM25
bypasses, and wall-clock, while keeping the main claim narrowly about LLM
calls and observed wall-clock.
