\section{Introduction}
\label{sec:introduction}

% Style note (v2, 2026-05-07): rewritten in the EMNLP 2025 Outstanding-paper
% style — concrete grounding example introduced before the method, contributions
% list tightened, two-question framing made explicit. Word target ~500 to fit
% the 4-page short-paper budget per plan v4.

Modern passage ranking is dominated by multi-stage pipelines in which a cheap first-stage retriever, often BM25, returns a candidate pool that is then reranked by a more expensive neural model~\citep{robertson2009probabilistic,nogueira2019passage,nogueira2020monot5,yates2021pretrained}. This setting underlies the standard benchmarks --- MS MARCO, the TREC Deep Learning Tracks, and BEIR --- which score rerankers by graded metrics such as nDCG@10~\citep{bajaj2016msmarco,craswell2020trecDL,craswell2021trecDL,thakur2021beir,jarvelin2002cumulated}. Recent zero-shot LLM rerankers revisit this pipeline by prompting rather than fine-tuning the model to judge, compare, or order passages~\citep{sachan2022upr,sun2023rankgpt,ma2023zeroshotlistwise,qin2024prp,zhuang2024setwise,pradeep2023rankvicuna,pradeep2023rankzephyr}.

LLM zero-shot reranking is organized around four prompting paradigms: pointwise scoring~\citep{nogueira2020monot5,sachan2022upr}, pairwise comparison~\citep{pradeep2021expandomonoduo,qin2024prp}, listwise permutation generation~\citep{ma2023zeroshotlistwise,sun2023rankgpt,pradeep2023rankvicuna,pradeep2023rankzephyr,yoon2024listt5}, and setwise selection~\citep{zhuang2024setwise}. The setwise interface is appealing because each call observes more than two documents but the model still emits a short label rather than a full permutation. Existing setwise rerankers, however, route this interface through a small candidate window of $c{+}1$ passages embedded inside a sorting procedure such as heapsort or bubblesort~\citep{qin2024prp,zhuang2024setwise}. The choice was natural under tight context budgets, and the same locality recurs in listwise sliding-window or tournament schedules~\citep{sun2023rankgpt,ma2023zeroshotlistwise,yoon2024listt5}.But it leaves a question that the original context-limited setting couldnot answer: \emph{when the live reranking pool already fits in context, how much of the windowed setwise machinery is still necessary?}

The context-length regime has changed. For first-stage reranking with $N\!\le\!100$, the query plus all candidate passages typically fit in a single prompt. With passage length $\ell{=}512$ and pool size $N{=}50$, a rendered prompt requires roughly 26K input tokens before output reserve, well below the long-context budgets of the Qwen3.5, Llama-3.1, and Ministral-3 families considered here~\citep{qwen2026qwen35,dubey2024llama3,liu2026ministral3}. This does not make windowed setwise reranking obsolete: long prompts can raise input-token cost and per-call latency, and long-context models do not necessarily attend uniformly across positions~\citep{liu2024lost,zheng2024mcqbias,shi2025positionbias,pradeep2023rankzephyr}. It does, however, motivate a missing baseline rather than a replacement.

We study this baseline through \textbf{MaxContext}, a zero-shot setwise reranker family whose every LLM call observes the current whole live pool rather than a fixed local window. Documents in the live pool $L$ are labeled $1,\ldots,|L|$, and the model is asked to return one or two extreme labels. \textsc{MaxContext-TopDown} requests the best document and places it at the head of the output ranking; \textsc{MaxContext-BottomUp} requests the worst document and places it at the tail; \textsc{MaxContext-DualEnd} requests both jointly in a single call, placing one at each end and removing two documents from $L$ per call. The two single-extreme variants close the residual two-document pair with a deterministic BM25 endgame.

This three-way design separates two questions that are easy to conflate. \emph{First}, is whole-pool setwise selection a viable alternative to windowed setwise sorting? \emph{Second}, does eliciting both ends of the ranking in the same call preserve effectiveness while reducing the number of serial LLM calls? Within MaxContext the call-count comparison is deterministic: for a pool of size $N$, \textsc{MaxContext-DualEnd} uses $\lfloor N/2 \rfloor$ LLM calls, whereas either single-extreme variant uses $N{-}2$ LLM calls plus the BM25 endgame. As a worked example: on $N{=}10$, \textsc{DualEnd} runs 5 calls vs.\ 8 for either single-extreme variant; on $N{=}100$, the gap widens to 50 vs.\ 98. We therefore test whether the joint best--worst interface delivers this serial-call reduction without a corresponding loss in nDCG@10.

We compare seven methods --- the four standard windowed setwise baselines of \citet{zhuang2024setwise} (TopDown / BottomUp $\times$ heapsort / bubblesort) and the three MaxContext variants --- across nine open-weight checkpoints from three families (Qwen3.5, Llama-3.1, Ministral-3), eight datasets, and pool sizes $N \in \{10,20,30,40,50,100\}$. Whole-pool prompting is not order-invariant, so we also instrument MaxContext with two per-comparison position-bias controls (reverse and fixed-seed shuffle of the live pool) to measure how much of any whole-pool result depends on candidate order rather than candidate content. We frame the efficiency claim narrowly --- \textsc{DualEnd} is designed to reduce serial LLM calls and we report wall-clock and tokens transparently but do not claim token efficiency~\citep{peng2025flopsreranking}.

Our contributions are threefold. (1) We introduce MaxContext, a whole-pool setwise reranker family with best-only, worst-only, and joint best--worst output policies, and we characterize its deterministic call-count gap as $\lfloor N/2 \rfloor$ versus $N{-}2$ within the family. (2) We provide a controlled cross-family comparison against standard windowed setwise rerankers across seven methods, nine models, eight datasets, and six pool sizes. (3) We evaluate effectiveness alongside input-order behavior under whole-pool prompting, instrumenting MaxContext with reverse and fixed-seed-shuffle controls so that any reordering effect is measurable rather than assumed away.
