\section{Results}
\label{sec:results}

% Style note (v2, 2026-05-07): scaffolded in the EMNLP 2025 Outstanding-paper
% Setup/Hypothesis/Results/Discussion micro-template. Each sub-section
% preregisters its hypothesis before reporting findings. Numerical
% [TBD] placeholders mark cells that the Phase B/C/F runs will populate;
% the rhetorical scaffold and statistical protocol are settled here so
% that camera-ready edits are local.

We organize the results around three questions, each tested over the same main matrix with the protocol of \S\ref{sec:experimental-setup}.

\subsection{Q1: Does whole-pool DualEnd preserve effectiveness while approximately halving serial LLM calls?}
\label{sec:results-cost-quality}

\paragraph{Setup.}
We compare the seven methods within each (model, dataset, pool-size) cell at the matched operating point $W{=}k{=}N$. The within-MaxContext contrast is between \textsc{DualEnd} and either single-extreme MaxContext variant, where call counts are deterministic (Table~\ref{tab:call-accounting}). The cross-family contrast is between the three MaxContext variants and the four standard windowed setwise baselines.

\paragraph{Hypothesis.}
We predict that \textsc{DualEnd} is non-inferior to
\textsc{MaxContext-TopDown} and \textsc{MaxContext-BottomUp} on per-query nDCG@10 at margin $\delta{=}0.01$, on at least 80\% of cells (the H3 threshold). We further predict that the three MaxContext variants are competitive with the standard windowed setwise baselines at the same matched operating point, since context-length is no longer the binding constraint that motivated the windowed design. We do \emph{not} predict that \textsc{DualEnd} dominates the field; the central efficiency claim is about serial LLM calls, not effectiveness.

\paragraph{Results.}
Table~\ref{tab:main-deltas} reports per-cell signed nDCG@10 differences $\Delta_{\mathrm{TD}} = \mathrm{nDCG@10}(\textsc{DualEnd}) - \mathrm{nDCG@10}(\textsc{MaxContext-TopDown})$ and $\Delta_{\mathrm{BU}} = \mathrm{nDCG@10}(\textsc{DualEnd}) - \mathrm{nDCG@10}(\textsc{MaxContext-BottomUp})$ at $N{=}50$, with $\checkmark$ / $\blacktriangle$ / $\times$ markers for the H1 non-inferiority verdict. [TBD: per-cell verdicts after Phase B completion.] Aggregated across the 4-model main-paper subset and 6 datasets, [TBD: $X/24$ cells favor non-inferiority for $\Delta_{\mathrm{TD}}$ and $Y/24$ for $\Delta_{\mathrm{BU}}$ under Holm-Bonferroni at $\alpha{=}0.05$]. Figure~\ref{fig:pareto} plots the mean LLM calls vs.\ mean nDCG@10 across the pool sweep $N \in \{10,20,30,40,50,100\}$.

\input{tables/main_deltas}
\input{figures/pareto}

\paragraph{Discussion.}
[TBD interpretation paragraph.] If the verdict pattern is supportive, this is direct evidence that joint best--worst elicitation amortizes the serial-call cost of whole-pool setwise selection. If supportive only below $N{=}100$, the discussion frames a long-context saturation point rather than a flat win. If the verdict is mixed, the call-count reduction remains exact (Table~\ref{tab:call-accounting}) but its practical value depends on the model family --- a discussion we resume in Q2.

\subsection{Q2: Does the result generalize across model families?}
\label{sec:results-family}

\paragraph{Setup.}
Each (method, dataset, pool-size) verdict from Q1 is broken out by model family --- five Qwen3.5 checkpoints, one Llama-3.1, three Ministral-3 --- and by parameter scale within Qwen3.5 (0.8B--27B). Tier-1 main-paper cells use four checkpoints chosen to vary family while holding scale near the 8B class; the remaining cells populate the appendix.

\paragraph{Hypothesis.}
We predict that the within-family DualEnd-vs-single-extreme contrast is qualitatively consistent across the three families, even where its magnitude is not. We do not predict that all families behave identically; differences in chat template, tokenizer, and long-context implementation are expected to produce family-specific effect sizes.

\paragraph{Results.}
[TBD per-family signed-$\Delta$ summary across the four Tier-1 checkpoints and six datasets at $N{=}50$, with per-family verdict counts.] Figure~\ref{fig:pareto} bottom panel reports the BEIR-mean Pareto curve per family. Family-stratified scale behavior within Qwen3.5 (0.8B--27B) is reported in the appendix.

\paragraph{Discussion.}
[TBD interpretation: whether the finding is a property of the algorithm or of a specific family.] The cross-family design is the principal guard against the single-family confound that motivated this study~(\S\ref{sec:methodology}); the specific pattern of agreement or disagreement across Qwen3.5, Llama-3.1, and Ministral-3 is the cleanest evidence we can report on family-generality.

\subsection{Q3: How sensitive is whole-pool MaxContext to candidate order?}
\label{sec:results-position-bias}

\paragraph{Setup.}
We pair each MaxContext-only forward run from Q1 with two per-comparison position-bias controls applied to the live pool $L$ before each LLM call: a deterministic \textbf{reverse} of $L$, and a fixed-seed shuffle of $L$ (seed 929). The required matrix covers three representative models, four representative datasets, three MaxContext methods, six pool sizes, and two control conditions ($3 \times 4 \times 3 \times 6 \times 2 = 432$ runs; \S\ref{sec:experimental-setup}). The analysis reports paired within-query nDCG@10 deltas for forward vs.\ reverse and forward vs.\ shuffle, plus selected-label position frequencies.

\paragraph{Hypothesis.}
We predict that whole-pool MaxContext is \emph{not} order-invariant: the forward-vs-reverse and forward-vs-shuffle deltas should be detectable above the deterministic-decoding noise floor, with magnitudes that correlate with pool size $N$ and may correlate with model family. We take this as a property of long-context LLM prompting~\citep{liu2024lostmiddle,tang2024foundmiddle,zheng2024mcqbias,shi2025positionbias,pradeep2023rankzephyr} rather than a critique of MaxContext.

\paragraph{Results.}
[TBD: per-cell paired delta distribution; aggregated summary statistics (mean, SD, range, worst-pair $|\Delta|$); selected-label position frequencies.] We expect the position-frequency view to expose whether selected-label distributions concentrate at specific live-pool positions under forward order and shift predictably under reverse and shuffle.

\paragraph{Discussion.}
[TBD interpretation.] If forward-vs-reverse deltas are large and systematic, this is a position-bias finding for whole-pool MaxContext that windowed setwise reranking does not surface --- the finding is specific to the long-context regime and does not refute the call-count result in Q1. If deltas are small, the long-context regime tolerates pool reordering and the call-count result is robust. The deterministic 10$\times$ stability check (\S\ref{sec:experimental-setup}) is reported separately as a byte-equality guard rather than as input-order evidence.
