Several limitations bound the scope of the claim.

\paragraph{Context fit is not attention quality.}
We evaluate $N \le 50$ at passage length $\ell=512$. This is an operating
point, not a universal limit: it follows from the chosen passage budget,
the model's configured context length, and a strict no-truncation
preflight. Larger pools may fit on longer-context deployments or shorter
passages, but fitting the prompt does not guarantee that the model uses
all positions equally well. The input-order protocol is therefore a
diagnostic, not proof of order invariance.

\paragraph{Call-count is not token-efficiency.}
MaxContext spends many more prompt tokens per call than small-window
setwise. Our headline comparison is restricted to serial LLM calls
($\lfloor N/2 \rfloor$ vs.\ $N{-}2$) and observed wall-clock. Prompt,
completion, and total-token costs are reported precisely because a reader
optimising for tokens may prefer a different point on the cost-quality
frontier.

\paragraph{Generation-only model scope.}
The main matrix covers Qwen3.5, Llama 3.1, and Ministral 3 under
generation scoring. Encoder-decoder backbones such as Flan-T5 and
likelihood-scoring paths are excluded because they do not model the joint
\texttt{Best: i, Worst: j} output directly; they collapse the dual
decision into independent best-only proxies and would test a different
mechanism.

\paragraph{Statistical power.}
TREC DL19 and DL20 contain only 43 and 54 judged queries. A per-cell
non-inferiority verdict is meaningful only relative to the predeclared
$\delta=0.01$ margin and the paired query distribution. We mitigate this
with paired-bootstrap CIs, BEIR domains with larger query counts, and
FDR-corrected aggregate verdicts, but individual cells remain noisy.

\paragraph{BM25-endgame asymmetry.}
The deterministic two-document endgame affects different rank positions:
TopDown resolves the tail ($N{-}1$ vs.\ $N$), while BottomUp resolves the
head (ranks $1$ vs.\ $2$). We therefore compare DualEnd separately
against each single-extreme control and avoid attributing any quality gap
solely to the joint prompt without a no-bypass diagnostic.

\paragraph{No primary listwise or tournament head-to-head.}
TourRank~\citep{chen2025tourrank}, BLITZRANK~\citep{blitzrank2026}, and
listwise rerankers such as RankGPT~\citep{sun2023rankgpt} use different
budget knobs, including grouping, advancement, graph queries, window
stride, and output length. A fair comparison requires matched-budget
sweeps over those knobs. This short paper instead isolates the
whole-pool setwise question and leaves broader head-to-head studies to
future work.
