Setwise reranking with large language models~\citep{zhuang2024setwise} prompts
the model to select the most relevant document from a small candidate window
and sorts these local decisions into a top-$k$ ranking, extracting only one
explicit extreme per LLM call --- the identity of the best document --- without
explicitly using the worst document or the relative order of the rest. Asking
\emph{jointly} for the best and worst document in the same prompt (DualEnd)
extracts two explicit extremes per call, but at the usual four-passage window
it still requires hundreds of dependent LLM calls per query, leaving a large
latency gap between lightweight setwise sorts and quality-oriented joint
elicitation.
We propose \textbf{MaxContext}, a family of zero-shot LLM rerankers that
prompts repeatedly over the \emph{whole live rerank pool} (up to $50$
documents) instead of a small sliding window. The family has three variants
that ask, on each call, for either joint best+worst (\textsc{DualEnd}),
best-only (\textsc{TopDown}), or worst-only (\textsc{BottomUp}) selection from
the live pool; \textsc{MaxContext-DualEnd} reranks a pool of size $N$ in
$\lfloor N/2 \rfloor$ dependent LLM calls, and the single-extreme variants in
$N{-}2$ calls plus a deterministic two-document BM25 endgame. The three
variants jointly test whether quality differences relative to small-window
setwise arise from joint elicitation, from whole-pool selection itself, or
from single-extreme selection at long context. We evaluate MaxContext on six
Qwen3 / Qwen3.5 checkpoints (4B--27B) on TREC~DL19 and DL20 against
matched-\texttt{hits} TopDown-Heap, TopDown-Bubble, DualEnd-Cocktail, and
DualEnd-Selection baselines, with explicit pool-size, passage-length, and
input-order ablations. As a secondary fixed-configuration operating point that can scale beyond
MaxContext's pool-size ceiling, we additionally evaluate a Top-$m$~Setwise
+ DualEnd finalist cascade. We frame
our claim as potential movement on the comparisons-axis and wall-clock-axis of
the reranking Pareto frontier with transparent reporting of prompt,
completion, and total tokens, parse-fallback rates, and order-stability
diagnostics, so that any quality difference can be audited against its true
compute cost rather than an isolated call-count headline.
